Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-10 ltx/paper.tex
a small change
received full performance metric dataset from Ellis/Ferryman, included bar graph of this data to show comparison of our technique with object-specific trackers

In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.

@@

\subsection{From Inference to Tracking Results}
\label{sec:inference_to_results}
The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results. The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}). In this study, we desired the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video. We chose to make our results depend completely on the inferred parameters of the multivariate normal distribution of each cluster. Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps. In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:pets2000_2001}) was taken to be the spatial region of the object. Furthermore, as both provided inference methods use sampling, and each sample represents a result, we must choose a way to narrow these down to a single result. We have found that calculating the joint log probability for each sample and choosing the sample that yields the highest value as a result works reasonably well.

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figure, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.

@@

The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}). In Figure \ref{fig:pets2009_results}, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.

\begin{figure}[!]
  \centering
  % \subfloat[]{\label{fig:pets2009_results} \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}}
  \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}
  \caption{Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
  \label{fig:pets2009_results}
  % \label{fig:benchmark_results_2}
\end{figure}

>>>>

In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of ``color counts'' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.

@@

\subsection{From Inference to Tracking Results}
\label{sec:inference_to_results}
The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results. The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}); in this study, we desire the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video. We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal distribution of each cluster. Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps. In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:pets2000_2001}) was taken to be the spatial region of the object. Furthermore, as both provided inference methods use sampling, and each sample represents a result, we must choose a way to narrow these down to a single result. We have found that calculating the joint log probability for each sample and choosing the sample that yields the highest value as a result works reasonably well.

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used from both datasets. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figures, a sample from the posterior distribution of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.

@@

The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}). In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.

\begin{figure}[!]
  \centering
  \subfloat[]{\includegraphics[width=0.43\textwidth]{../img/pets2009_1-50_front.pdf}}
  \hspace{4mm}
  \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b.pdf}}
  \caption{ (a) Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. (b) performance metrics vs. covariance confidence interval threshold for tracking results showing the SFDA (blue solid line) and ATA (red dashed line).}
  \label{fig:pets2009_results}
\end{figure}

In \cite{ellis_2010}, performance metrics were computed for a number of studies that carried out object detection and tracking for the PETS2009/2010 dataset. As this dataset consists solely of humans, many of these algorithms were developed for the specific purpose of people tracking (i.e. not for general detection and tracking of arbitrary objects); in consequence, many of these studies involve appearance models specifically developed for humans, sometimes making use of their orientation in this specific dataset, and motion models based on assumptions about human motion. We compare SFDA and ATA results of our strategy with these methods to show that our general method can yield comparable results even when compared with object-specific trackers. The following table shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
\begin{figure}[!]
  \centering
  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_sfda.pdf}}
  \hspace{1mm}
  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_ata.pdf}}
  \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops shows comparable performance with state-of-the-art, object-specific detection and tracking strategies. Results for the technique in this paper are shown in red, and were computed using a covariance confidence interval threshold of 0.35.}
  \label{fig:sfda_and_ata_comparison}
\end{figure}
